---
title: "Week3"
author: "Aybike Dilek"
date: "04 11 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
```

## In-Class Exercise

The data set contains various data on the housing sector in Istanbul. Data are available from 2010 to 2020 on a monthly basis.


```{r}
data <- read_excel("C:/Users/Aybike/Desktop/data.xlsx")
summary(data)
colnames(data)[2]="TP_AKONUTSAT1_T40"
colnames(data)[6]="TP_DISKONSAT_ISTANBUL"
summary(data)
sub_data<-data %>% select(Tarih,TP_AKONUTSAT1_T40,TP_DISKONSAT_ISTANBUL) %>% mutate(Gün=as.Date(paste0(Tarih,"-01"))) %>% mutate(Yıl=year(Gün))
summary(sub_data)
tsp <- ggplot(data = sub_data, aes(x = Yıl, y = TP_AKONUTSAT1_T40 ), y_lim=seq(2013,2020,1))+ geom_line(color = "#00AFBB", size = 1)
tsp

```


I observed that home sales increased sharply after 2018. I think that it could be due to housing phase discount and urban transformation.

```{r}
tsp <- ggplot(data = sub_data, aes(x = Yıl, y = TP_DISKONSAT_ISTANBUL ), y_lim=seq(2013,2020,1))+ geom_line(color = "#00AFBB", size = 1)
tsp

```

I observed that home sales to foreigners increased sharply after 2017. I think that it could be due to appreciation of Dollar currency against Turkish Lira.