---
title: "Week3 In-Class Exercise"
author: "Aybike Dilek"
date: "*04/11/2020*"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: united
    highlight: pygments
  
---
```{css, echo = FALSE}
h1.title{
  color:#d3501d;
}
h2{
  color:#084a4a;
}
h3{
  color:#084a4a;
}
h4{
  color:#084a4a;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(scales)
```

## In-Class Exercise

The data set contains various data on the housing sector in Istanbul. Data are available from 2010 to 2020 on a monthly basis.

### Preprocessing

```{r}
data <- read_excel("C:/Users/Aybike/Desktop/data.xlsx")
summary(data)
colnames(data)[5]="TP_AKONUTSAT4_T40"
colnames(data)[6]="TP_DISKONSAT_ISTANBUL"
summary(data)
sub_data<-data %>% select(Tarih,TP_AKONUTSAT4_T40,TP_DISKONSAT_ISTANBUL) %>% mutate(Gün=as.Date(paste0(Tarih,"-01"))) %>% mutate(Yıl=year(Gün)) 
sum_data<- sub_data %>% group_by(Yıl) %>% summarise(Toplam_TP_AKONUTSAT4_T40 =  sum(TP_AKONUTSAT4_T40),Toplam_TP_DISKONSAT_ISTANBUL=sum(TP_DISKONSAT_ISTANBUL))
sum_data
```

### Data Analysis with R  

```{r}
tsp <- ggplot(data = sum_data, aes(x = Yıl, y = Toplam_TP_AKONUTSAT4_T40))+geom_line(color="#d3501d",size=1)+
labs(title = "Toplam_TP_AKONUTSAT4_T40 vs Yıl")+
xlim(2013,2019)
tsp
```

*Toplam_TP_AKONUTSAT4_T40 = Sum of second-hand house sales in ISTANBUL*

*Yıl = Year*

I observed that second-hand house sales significantly increased in 2019. I think that it could be due to decreasing of affordability.

* 2020 is an incomplete year so it is excluded.

```{r}
tsp <- ggplot(data = sum_data, aes(x = Yıl, y = Toplam_TP_DISKONSAT_ISTANBUL))+ geom_line(color = "#00AFBB", size = 1) + xlim(2013,2019)+labs(title = "Toplam_TP_DISKONSAT_ISTANBUL vs Yıl")
tsp
```

*Toplam_TP_DISKONSAT_ISTANBUL= Sum of house sales to foreigners in ISTANBUL*

*Yıl = Year*

I observed that home sales to foreigners increased sharply after 2017. I think that it could be due to appreciation of Dollar currency against Turkish Lira.

* 2020 is an incomplete year so it is excluded.

### References

* [Data Source-1](http://www.tuik.gov.tr/)

* [Data Source-2](http://www.tcmb.gov.tr/)

*You may click* [here](https://pjournal.github.io/mef04-AybikeDilek/) *to reach other items of my progress journal.* 